---
title: "Precipitation Interpolation"
author: "Matthew Ross"
date: "4/13/2022"
output: html_document
---

```{r setup, include=FALSE}

library(tidyverse)
library(sf) # new goodness
library(mapview)
library(lubridate)
library(osmdata)
library(gstat)
library(elevatr)
library(USAboundaries)
library(leafpop) #map html popup
library(stars)
library(viridis)


knitr::opts_chunk$set(echo = TRUE)
```


# Load in Precip data

```{r, warning=F, message=F, results='hide'}
load('data/DailyP.RData')

daily_p %>% glimpse()
```


## Get Elevation Data

```{r, eval = T}
if(file.exists("data/unique_asos_elev.gpkg") == FALSE){
  unique_asos <- daily_p %>%
    distinct(lon, lat, station)  %>%
    st_as_sf(., coords = c('lon','lat'), crs = 4326) %>%
    get_elev_point(.)
  
  st_write(unique_asos, "data/unique_asos_elev.gpkg")
}
```

## Read elevation data
```{r, eval = TRUE}
unique_asos <- st_read("data/unique_asos_elev.gpkg")
```

## Get Monthly P Averages

```{r, warning=F, message=F, results='hide'}
monthly_p <- daily_p %>%
  mutate(month = month(date)) %>%
  group_by(month, station) %>%
  summarize(monthly_p = sum(daily_p)) %>%
  left_join(unique_asos, .) #grab elevation data
```


## Look at monthly P

```{r, warning=F, message=F, fig.width = 8, fig.height = 5}
ggplot(monthly_p, aes(x = elevation, y = monthly_p, color = month)) + 
  scale_color_viridis_c() + 
  geom_point()
```


## Getting Monthly Means of means, mins, maxes. 

```{r, warning=F, message=F, results='hide'}
monthly_t <- daily_p %>%
  mutate(month = month(date)) %>%
  group_by(month, station) %>%
  select(-lon,-lat) %>%
  summarize(across(where(is.numeric), mean, na.rm = T)) %>%
  left_join(unique_asos, .)
```


## Temp vs Elevation

```{r, warning=F, message=F, fig.width = 8, fig.height = 5}
ggplot(monthly_t, aes(y = mean_temp, x = elevation, color = month)) + 
  geom_point() + 
  scale_color_viridis_c()
```

### Set up CO state spatial

```{r, warning=F, message=F, results='hide'}
# get state polygon
co_state <- us_states(states = c("Colorado")) %>% 
  st_transform(2163)

# make grid out of state bounding box and crop to state
co_state_stars <- st_bbox(co_state) %>%
  st_as_stars(dx = 1000) %>%
  st_crop(co_state)

# change mapview defaults
mapviewOptions(homebutton = FALSE, basemaps = c("Esri"))
```

## Pick a month (summer months are safer)

```{r, warning=F, message=F, results='hide'}
my_idw_fn <- function(my_month = 7, response_v = "daily_p", predictor = 1, lab = "Mean Precipitation"){
# filter for month
  pushin_p <- monthly_t %>% 
    filter(month == my_month) %>% 
    mutate(across(where(is.numeric), ~ ifelse(is.infinite(.x), NA, .x)
                  )) %>% 
    filter(if_all(starts_with(response_v), ~ !is.na(.))) %>% # gets rid of Inf and NA so IDW works
    st_transform(2163)
# Build IDW
  fmla <- as.formula(paste(response_v, "~", predictor))
  idw_intrpltn <- idw(fmla, pushin_p, co_state_stars)
# Make plot
  idw_map <- mapview(idw_intrpltn
        , layer.name = lab
        , col.regions= mapviewGetOption('vector.palette')
        , popup = FALSE
  ) + mapview(pushin_p
        , zcol = response_v
        , legend = FALSE
  )
return(list(idw_map, idw_intrpltn))
}
```


### Build IDW precip or elevation for state for that month

```{r, warning=F, message=F, results='hide'}
rslts <- my_idw_fn(my_month = 8, response_v = "max_temp", predictor = 1, lab = "Aug Max Temp")
```

### Plot this data

```{r, warning=F, message=F, fig.width = 8, fig.height = 5}
rslts[[1]]
```

### Build IDW with elevation for state for that month including elevation as a predictor

Hint! Use `get_elev_raster` use a z of 7 or 8

```{r, warning=F, message=F, results='hide'}
elev_stars <- get_elev_raster(co_state, z = 7) %>% 
  st_transform(2163)

```

### Make a Map of that

```{r, warning=F, message=F, fig.width = 8, fig.height = 5}
## You will need to create a Stars raster that has elevation data. 



```


### Compare both maps to PRISM approach for your month


How close do our simple approaches come to reproducing prism maps? 


https://www.prism.oregonstate.edu/recent/monthly.php


```{r, warning=F, message=F, fig.width = 8, fig.height = 5}

```

