---
title: "Precipitation Interpolation"
author: "Matthew Ross | Assignment completed by: George Woolsey"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
knit: (function(inputFile, encoding){ 
    out_dir <- '/';
    rmarkdown::render(inputFile, encoding = encoding, output_file=file.path(dirname(inputFile), out_dir, 'index.html')) 
  })
---

```{r setup, include=FALSE}

library(tidyverse)
library(sf) # new goodness
library(mapview)
library(lubridate)
library(osmdata)
library(gstat)
library(elevatr)
library(USAboundaries)
library(leafpop) #map html popup
library(stars)
library(viridis)
library(tmap) # make publication quality maps
library(starsExtra)


knitr::opts_chunk$set(echo = TRUE)
```


# Classwork

## Load Precip. data

```{r, warning=F, message=F, results='hide'}
load('data/DailyP.RData')

daily_p %>% glimpse()
```


## Get Elevation Data

```{r, eval = T}
if(file.exists("data/unique_asos_elev.gpkg") == FALSE){
  unique_asos <- daily_p %>%
    distinct(lon, lat, station)  %>%
    st_as_sf(., coords = c('lon','lat'), crs = 4326) %>%
    get_elev_point(.)
  
  st_write(unique_asos, "data/unique_asos_elev.gpkg")
}
```

## Read elevation data
```{r, eval = TRUE}
unique_asos <- st_read("data/unique_asos_elev.gpkg")
```

## Get Monthly P Averages

```{r, warning=F, message=F, results='hide'}
monthly_p <- daily_p %>%
  mutate(month = month(date)) %>%
  group_by(month, station) %>%
  summarize(monthly_p = sum(daily_p)) %>%
  left_join(unique_asos, .) #grab elevation data
```


## Look at monthly P

```{r, warning=F, message=F, fig.width = 8, fig.height = 5}
ggplot(monthly_p, aes(x = elevation, y = monthly_p, color = month)) + 
  scale_color_viridis_c() + 
  geom_point()
```


## Getting Monthly means and sums of measures

```{r, warning=F, message=F, results='hide'}
monthly_t <- daily_p %>%
  mutate(month = month(date)) %>%
  group_by(month, station) %>%
  select(-lon,-lat) %>%
  summarize(across(where(is.numeric), mean, na.rm = T, .names = "{.col}_mean")
            , monthly_p_sum = sum(daily_p)
      ) %>%
  # summarize(across(where(is.numeric), list(mean = mean, sum = sum), na.rm = T)) %>%
  left_join(unique_asos, ., by = c("station"="station"))
```


## Temp vs Elevation

```{r, warning=F, message=F, fig.width = 8, fig.height = 5}
ggplot(monthly_t, aes(y = mean_temp_mean, x = elevation, color = month)) + 
  geom_point() + 
  scale_color_viridis_c()
```

# Assignment

## Set up CO state spatial

```{r, warning=F, message=F, results='hide'}
# get state polygon
co_state <- us_states(states = c("Colorado")) %>% 
  st_transform(2163)

# make grid out of state bounding box and crop to state
co_state_stars <- st_bbox(co_state) %>%
  st_as_stars(dx = 1000) %>% # 1km grid
  st_crop(co_state)

# change mapview defaults
mapviewOptions(homebutton = FALSE, basemaps = c("Esri"))
```

## Build a function for IDW model

```{r, warning=F, message=F, results='hide'}
my_idw_fn <- function(
    my_month = 7
    , response_v = "monthly_p_sum"
    , predictor = c("1")
    , lab = "July Total Precipitation"
    , col_drctn = 1
    , viridis_pal_opt = "G"
  ){
# filter for month
  pushin_p <- monthly_t %>% 
    filter(month == my_month) %>% 
    mutate(across(where(is.numeric), ~ ifelse(is.infinite(.x), NA, .x)
                  )) %>% 
    filter(if_all(starts_with(response_v), ~ !is.na(.))) %>% # gets rid of Inf and NA so IDW works
    st_transform(2163)
# Build IDW
  fmla <- as.formula(paste(response_v, paste(predictor, collapse=" + "), sep=" ~ "))
  idw_intrpltn <- gstat::idw(fmla, pushin_p, co_state_stars)
# Make plot
  idw_map <- mapview(idw_intrpltn
        , layer.name = lab
        , col.regions= viridis_pal(direction = col_drctn, option = viridis_pal_opt)
        , na.color = NA
        , popup = FALSE
  ) + mapview(pushin_p
        , zcol = response_v
        , col.regions= viridis_pal(direction = col_drctn, option = viridis_pal_opt)
        , legend = FALSE
  )
return(list(idw_map, idw_intrpltn))
}
```


```{r, include=FALSE}
# unique_asos_t <- st_transform(unique_asos, crs = 2163)
# co_box <- st_bbox(unique_asos_t) %>% 
#   st_as_stars(dx=1000)
# 
# 
# july_t <- monthly_t %>% 
#   dplyr::filter(month == 7) %>% 
#   st_transform(., st_crs(co_box))
# 
# interp <- idw(mean_temp~1, july_t, co_box)
# 
# plot(interp)
# mapview(interp)
# 
# tm_shape(interp[1]) + tm_raster(palette = 'Reds', style = 'cont')
# 
# #elevation impacted model of temp
# ras <- get_elev_raster(unique_asos, z=8) %>% 
#   raster::crop(., unique_asos)
# 
# mapview(ras)
# co_stars <- st_as_stars(ras)
# 
# names(co_stars) <- 'elevation'
# 
# june_t <- monthly_t %>% 
#   filter(month == 6)
# 
# interp = gstat::idw(mean_temp~elevation, june_t, co_stars)
# 
# tm_shape(interp[1]) + tm_raster(palette = 'plasma', style='cont')
# 

```

## Ordinary and simple IDW

```{r, warning=F, message=F, results='hide'}
# total precip
precip_rslts <- my_idw_fn(my_month = 8, response_v = "monthly_p_sum"
                          , predictor = 1, lab = "Aug Total Precip"
                          , col_drctn = -1, viridis_pal_opt = "G")
# avg max temp
maxt_rslts <- my_idw_fn(my_month = 8, response_v = "max_temp_mean"
                          , predictor = 1, lab = "Aug Avg Max Temp (F)"
                          , col_drctn = 1, viridis_pal_opt = "E")

```

<span style="color: blue;">Above, the ordinary and simple IDW model uses the formula `y~1`. Since no independent variables are defined the model only contains an intercept and beta should be the simple IDW mean. </span>

### Plot this data

```{r, warning=F, message=F, fig.width = 8, fig.height = 5}
# precipitation IDW
# spatial interpolation only
precip_rslts[[1]]

# average max temp IDW
# spatial interpolation only
maxt_rslts[[1]]
```

<span style="color: blue;">The map above uses `1` as a predictor of temperature/precipitation which means that the IDW model uses spatial interpolation only. Thus, there is a lack of predictive power in areas where there is no station data collected. </span>

## Beyond ordinary and simple interpolation

Build IDW with elevation for the state to predict temperature for that month

Hint! Use `get_elev_raster` use a z of 6 or 7 or 8

### Linear trend elevation and temperature

The universal IDW supposes that `y` is linearly dependent on `x1` and `x2` and thus uses the formula `y~x1+x2`.

```{r, warning=F, message=F, fig.width = 8, fig.height = 5}

ggplot(monthly_t %>% filter(month == 8), aes(x = elevation, y = max_temp_mean)) +
  geom_point(alpha=0.8, color = "navy", size = 2) + 
  geom_smooth(method = 'lm', color = "grey35", alpha = 0.8) +
  labs(
    title = "Relationship between elevation and Avg. Max Temperature (F)"
    , subtitle = "August 2019"
  ) +
  xlab("Elevation (m)") +
  ylab("Avg. Max Temp. (F)") +
  theme_bw() +
  theme(
    legend.position="none"
  )
  

```

### Get elevation grid

```{r, warning=F, message=F, results='hide'}
elev_raster <- get_elev_raster(co_state, z = 6)
names(elev_raster)[1] <- c("elevation")

co_state_stars <- elev_raster %>% 
  st_as_stars(.) %>% 
  st_crop(co_state)
```

### Build Model

```{r, warning=F, message=F, results='hide'}
maxt_elev <- my_idw_fn(my_month = 8, response_v = "max_temp_mean"
                          , predictor = "elevation", lab = "Aug Avg Max Temp (F)"
                          , col_drctn = 1, viridis_pal_opt = "E")

```

### Make a Map of that

```{r, warning=F, message=F, fig.width = 8, fig.height = 5}
maxt_elev[[1]]
```

<span style="color: blue;">Note, the increased predictive resolution when including variation in both elevation and geographic space as predictors. The universal IDW supposes that `y` is linearly dependent on `x1` and `x2` and thus uses the formula `y~x1+x2`.</span>

### Compare both maps to PRISM approach for your month

How close do our simple approaches come to reproducing prism maps? 

[PRISM maps](https://www.prism.oregonstate.edu/recent/monthly.php)

<span style="color: blue;">The map created above using elevation as a predictor of temperature aligns well with the PRISM map and improves upon the spatial interpolation only mo.</span>

## Elevation & aspect as predictors

Build IDW with elevation and aspect for the state to predict temperature for that month

### Calculate aspect

```{r, warning=F, message=F, fig.width = 8, fig.height = 5}
# calculate aspect
aspect <- starsExtra::aspect(co_state_stars[1])
names(aspect)[1] <- c("aspect")

# attach aspect to original stars data
co_state_stars <- c(co_state_stars, aspect) %>% 
  mutate(cos_aspect = cos(aspect))

# quick plot
ggplot() +
  geom_stars(data = aspect %>% st_transform(., crs=st_crs(unique_asos))) +
  # geom_stars(data = co_state_stars[3] %>% st_transform(., crs=st_crs(unique_asos))) +
  geom_sf(data = unique_asos) +
  theme_bw()

# join aspect to station data
asos_aspect <- st_extract(
                    aspect %>% st_warp(., crs = st_crs(unique_asos))
                    , unique_asos
              ) %>% 
  st_as_sf()

# set up table to join to monthly data
join_aspect <- unique_asos %>% 
  st_join(., asos_aspect) %>% 
  select(station, aspect) %>% 
  st_set_geometry(., NULL) %>% 
  as_tibble() 

# join to monthly data used in IDW
monthly_t <- monthly_t %>% 
  left_join(., join_aspect, by = c("station"="station"))
```

### Build model

```{r, warning=F, message=F, results='hide'}
maxt_elev_aspect <- my_idw_fn(my_month = 8, response_v = "max_temp_mean"
                          , predictor = c("elevation", "aspect"), lab = "Aug Avg Max Temp (F)"
                          , col_drctn = 1, viridis_pal_opt = "E")
```

### Make a Map of that

```{r, warning=F, message=F, fig.width = 8, fig.height = 5}
maxt_elev_aspect[[1]]
```

<span style="color: blue;">The map created above uses both elevation and aspect (in addition to spatial interpolation) as a predictor of temperature.</span>
